EGGs are adhesive!

### 2025-06-15 calco/presentazione/pres.tex
minori

\documentclass{beamer}
\usepackage{beamer}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{ragged2e}
\usepackage{xcolor}
\colorlet{shaded}{gray!40}

@@

    \emph{e-graphs (EGGS)} \cite{DetlefsNS05,WillseyNWFTP21} are a graphical formalism  for program optimisation  via an efficient implementation of equality saturation.

@@

\begin{frame}{$\mathcal{M}$-adhesive categories}
	The categorical framework for DPO rewriting is given by \emph{$\mathcal{M}$-adhesive categories} \cite{azzi2019essence,ehrig2014adhesive, lack2005adhesive}.

@@

		\parbox{5cm}{\centering \begin{tikzpicture}
				\node(C)at(-1,0.4){$C$};
				\node(A)at(1,0.8){$A$};
				\node(B)at(2,0.4){$B$};
				\node(D)at(0,0){$D$};
				\node(A')at(1,2){$A'$};
				\node(B')at(2,1.6){$B'$};
				\node(C')at(-1,1.6){$C'$};
				\node(D')at(0,1.2){$D'$};
				\draw[<-](B')--(A');
				\draw[>->](B')--(B);
				\draw[>->](C')--(C); 
				\draw[>->](D')--(D);
				\draw[->](A')--(C'); 
				\draw[->](B')--(D');
				\draw[->](C')--(D');
				\draw[->](A)--(B);
				\draw[->](C)--(D);
				\draw[->](B)--(D);
				\draw[>-](A')--(1,1.45);
				\draw[->](1,1.35)--(A);
				\draw[>-](A)--(0.05, 0.61);
				\draw[<-](C)--(-0.05,0.59);
		\end{tikzpicture}} \quad 	\parbox{4cm}{\centering
			Top face is a pushout \\ $\iff$ \\ Front faces are pullbacks}

@@

A hypergraph is a 4-uple $(E, V, s, t)$ made by two sets $E$ and $V$ (\emph{edges} and \emph{nodes}) and arrows $s, t\colon E \rightrightarrows V^\star$. \pause A \emph{hypergraph morphism} $(E, V, s, t)\to (F, W, s,' t')$  is a pair of arrow $h\colon E\to E'$ and $k\colon V\to V'$ fitting in the digram below.

@@

	\pause  \medskip 
	\begin{minipage}[l]{0.45\linewidth}
		Limits are more difficult. In the example on the right, the pullback of the underlying hypergraphs gives us the empty one. The pullback of quotients is the singleton.
	\end{minipage} \qquad 
\begin{minipage}[r]{0.6\linewidth}
	\xymatrix{
		 &
		\begin{tikzpicture}[baseline=(v.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v) at (0, 0){};
				\node[style=none] at(0, 0.5){$c$};
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}\draw[dashed, rounded corners] (-0.3, -0.3) rectangle (0.3, 0.3);\end{pgfonlayer}
		\end{tikzpicture}  \ar@{>->}@(d,)[d]!<0ex,0ex> \\ 
		\begin{tikzpicture}[baseline=(v.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v)at(0, 0){};
				\node[style=none]at(0, 0.5){$b$};
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}\draw[dashed, rounded corners] (-0.3, -0.3) rectangle (0.3, 0.3);\end{pgfonlayer}
		\end{tikzpicture}
		\ar@{>->}@(r,)@<-.5ex>[r] & 
		\begin{tikzpicture}[baseline=(v.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v)at(-0.5, 0){};
				\node[style=none]at(-0.5, 0.5){$b$};
				\node[style=node]at(0.5, 0){};
				\node[style=none]at(0.5, 0.5){$c$};
			\end{pgfonlayer}\begin{pgfonlayer}{eqlayer}
				\draw[dashed, rounded corners](-0.8, 0.3) rectangle (0.8, -0.3);
		\end{pgfonlayer}\end{tikzpicture}
		 }
	\end{minipage}
\end{frame}

@@

\medskip How far can we move way from $\catname{Set}$?\pause 
\begin{itemize}
	\item Edges and nodes can be taken from a suitable category.   
	\item Combine various categorical construction to add features to EGGs (e.g.~hierarchy of edges) while retaining adhesivity.
\end{itemize}
\end{frame}

>>>>

\documentclass[aspectratio=169]{beamer}
\usepackage{beamer}
\usepackage{amsthm,amsmath,amssymb,mathrsfs, dsfont}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{ragged2e}
\usepackage{xcolor}
\colorlet{shaded}{gray!40}

@@

    \emph{E-graphs (EGGS)} \cite{DetlefsNS05,WillseyNWFTP21} are a graphical formalism  for program optimisation  via an efficient implementation of equality saturation.

@@

\begin{frame}{$\mathcal{M}$-adhesive categories}
	A suitable categorical framework for DPO rewriting is given by the notion of \emph{$\mathcal{M}$-adhesive categories} \cite{azzi2019essence,ehrig2014adhesive, lack2005adhesive}.

@@

		\parbox{5cm}{\centering \begin{tikzpicture}
				\node(C)at(-1,0.4){$C$};
				\node(A)at(1,0.8){$A$};
				\node(B)at(2,0.4){$B$};
				\node(D)at(0,0){$D$};
				\node(A')at(1,2){$A'$};
				\node(B')at(2,1.6){$B'$};
				\node(C')at(-1,1.6){$C'$};
				\node(D')at(0,1.2){$D'$};
				\draw[<-](B')--(A');
				\draw[>->](B')--(B);
				\draw[>->](C')--(C); 
				\draw[>->](D')--(D);
				\draw[->](A')--(C'); 
				\draw[->](B')--(D');
				\draw[->](C')--(D');
				\draw[->](A)--(B);
				\draw[->](C)--(D);
				\draw[->](B)--(D);
				\draw[>-](A')--(1,1.45);
				\draw[->](1,1.35)--(A);
				\draw[>-](A)--(0.05, 0.61);
				\draw[<-](C)--(-0.05,0.59);
		\end{tikzpicture}} \qquad 	\parbox{4cm}{\centering
			Top face is a pushout \\ $\iff$ \\ Front faces are pullbacks}

@@

A \emph{hypergraph} is a 4-uple $(E, V, s, t)$ made by two sets $E$ and $V$ (\emph{edges} and \emph{nodes}) and arrows $s, t\colon E \rightrightarrows V^\star$. \pause A \emph{hypergraph morphism} $(E, V, s, t)\to (F, W, s,' t')$  is a pair of arrow $h\colon E\to E'$ and $k\colon V\to V'$ fitting in the digram below.

@@

	\pause  \medskip 
	\begin{minipage}[l]{0.5\linewidth}
		Limits are more difficult. In the example on the right, the pullback of the underlying hypergraphs gives us the empty one. The pullback of quotients is the singleton.
	\end{minipage} \qquad \quad 
\begin{minipage}[r]{0.6\linewidth}
	\xymatrix{
		 &
		\begin{tikzpicture}[baseline=(v.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v) at (0, 0){};
				\node[style=none] at(0, 0.5){$c$};
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}\draw[dashed, rounded corners] (-0.3, -0.3) rectangle (0.3, 0.3);\end{pgfonlayer}
		\end{tikzpicture}  \ar@{>->}@(d,)[d]!<0ex,0ex> \\ 
		\begin{tikzpicture}[baseline=(v.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v)at(0, 0){};
				\node[style=none]at(0, 0.5){$b$};
			\end{pgfonlayer}
			\begin{pgfonlayer}{eqlayer}\draw[dashed, rounded corners] (-0.3, -0.3) rectangle (0.3, 0.3);\end{pgfonlayer}
		\end{tikzpicture}
		\ar@{>->}@(r,)@<-.5ex>[r] & 
		\begin{tikzpicture}[baseline=(v.base)]\begin{pgfonlayer}{nodelayer}
				\node[style=node](v)at(-0.5, 0){};
				\node[style=none]at(-0.5, 0.5){$b$};
				\node[style=node]at(0.5, 0){};
				\node[style=none]at(0.5, 0.5){$c$};
			\end{pgfonlayer}\begin{pgfonlayer}{eqlayer}
				\draw[dashed, rounded corners](-0.8, 0.3) rectangle (0.8, -0.3);
		\end{pgfonlayer}\end{tikzpicture}
		 }
	\end{minipage}
\end{frame}

@@

\medskip How far can we move way from $\catname{Set}$?\pause 
\begin{itemize}
	\item Edges and nodes can be taken from another suitable category.   \pause 
	\item Combine various categorical construction to add features to EGGs (e.g.~hierarchy of edges) while retaining adhesivity.
\end{itemize}
\end{frame}
